Timely Object Recognition

### 2012-11-03 writeups/nips12/technical.tex
camera-ready draft sent to collaborators

As shown in Figure~\ref{fig:figure1}, the system is given two times: the setup time $T_s$ and deadline $T_d$.
From the setup time to the deadline, we want to obtain the best possible answer if stopped at any given time.
A single-number metric that corresponds to this objective is the area captured under the curve between the start and deadline bounds, normalized by the total area.
We evaluate policies by this more robust metric and not simply by the final performance at deadline time for the same reason that Average Precision is used instead of a fixed Precision vs. Recall point in the conventional evaluations.

\subsection{Sequential Execution}
An \emph{open-loop} policy takes actions in a sequence that does not depend on observations received from previous actions.
The common classifier cascade is an example \cite{Viola2001}.
In contrast, our goal is to learn a dynamic, or \emph{closed-loop}, policy, which would exploit the signal in scene and inter-object context for a maximally efficient path through the actions.

We refer to the information available to the decision process as the \emph{state} $s$.
The state includes the currently believed distribution over class presence variables $P(\mathbf{C}) = \{P(C_0), \ldots, P(C_K)\}$, where we write $P(C_k)$ to mean $P(C_k=1)$.

@@

The specific actions we consider in the following exposition are detector actions $a_{{det}_i}$, where ${det}_i$ is a detector class $C_i$, and a scene-level context action $a_{gist}$, which updates the probabilities of all classes.

\subsection{Selecting actions} \label{sec:value}
As our goal is to pick actions dynamically, we want a function $Q(s,a): S \times \mathcal{A} \mapsto \mathbb{R}$, where $S$ is the space of all possible states, to assign a value to a potential action $a \in \mathcal{A}$ given the current state $s$ of the decision process.
We can then define the desired policy $\pi$ as simply taking the (untaken) action with the maximum value:
\begin{align}
\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} Q(s,a_i)
\end{align}

Although the action space $\mathcal{A}$ is quite manageable, consisting of the detectors and global classifier we would like to run on the image, the space of possible states $S$ is intractable.
Therefore we cannot learn a tabular representation of $Q(s,a)$, and must use function approximation to represent it \cite{Sutton1998}.
We featurize the state-action pair and assume linear structure: $Q^\pi(s,a_i) = \theta_\pi^\top  \phi(s,a_i)$.

The policy's performance at time $t$ is determined by the detections that are part of the set of observations $\mathbf{o}^j$ at the last state $s^j$ before $t$.
Therefore, the final AP vs. Time evaluation of an episode is a function $eval(h,T_s,T_d)$ of the history of execution $h=s^0,s^1,\dots,s^J$.
It is precisely the normalized area under the AP vs. Time curve between $T_s$ and $T_d$, as determined by the detections in $\mathbf{o}^j$ for all steps $j$ in the episode.

As shown in \autoref{fig:rewards}, this evaluation function is additive per action, as each action can generate detections that either raise or lower the mean AP of the results so far ($\Delta ap$) and takes a certain time ($\Delta t$).
From these and $T_s$ and $T_d$, we can find the area under the curve that was contributed by the action.
So, we can then represent the final evaluation $eval(h,T_s,T_d)$ in terms of individual rewards: $\sum_{j=0}^J R(s^j,a^j)$.

Specifically, as shown in Figure~\ref{fig:rewards}, we define the \emph{reward} of an action as
\begin{align}\label{eq:advanced}
R(s^j,a_i) = \Delta \text{ap}_i (t_T^j-\frac{1}{2}\Delta t_i)
\end{align}
where $t_T^j$ and $\text{ap}^j$ are the time left until the deadline and the AP at state $s^j$, and $\Delta t_i$ and $\Delta \text{ap}_i$ are the time taken and AP change produced by the action $a_i$.
(For clarity of exposition, we do not account for $T_s$ here.)

\subsection{Learning the policy}
The expected value of the final evaluation can be written recursively in terms of the value function:
\begin{align} \label{eq:recursive_value}
Q^\pi(s^j,a_i) = \mathbb{E}_{s^{j+1}} [R(s^j,a_i) + \gamma Q^\pi(s^{j+1},\pi(s^{j+1}))]
\end{align}
where $\gamma \in [0,1]$ is a \emph{discount} value that can mitigate the effects of increasing state-transition uncertainty over long episodes.

@@

Learning the policy is then a problem of repeatedly gathering samples with the current policy, minimizing the error between the discounted reward to the end of the episode as predicted by our current $Q(s^j,a_i)$ and the actual values gathered, and updating the policy with the resulting weights.
This is fitted Q-iteration, a variant of generalized policy iteration \cite{Ernst2005,Sutton1998}.

@@

We use $L_2$-regularized regression to minimize the error.
We run $15$ iterations of accumulating samples by running $350$ episodes, starting with a baseline policy which will be described in \autoref{sec:evaluation}, and cross-validating the regularization parameter at each iteration; samples are not thrown away between iterations.

A meta-parameter of the approach is the discount $\gamma$.
With $\gamma=0$, the value function is determined entirely by the immediate reward.
Learning in this case can only result in completely greedy policies.

With $\gamma=1$, the value function is determined by the expected rewards to the end of the episode, and so should be a close approximation to the final evaluation metric.
However, the highest value for $\gamma$ is not necessarily best if the action-value function is not expressive enough to represent the actual state transition behavior of the world.
We experiment with several values of $\gamma$, and find a mid-level value ($0.4$) to work best.

\subsection{Feature representation}
Our policy is at its base determined by a linear function of the features of the state: $\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \theta_\pi^\top \phi(s,a_i)$.
Since we want to be able to learn a dynamic policy, the observations $\mathbf{o}$ that are part of the state $s$ should play a role in determining the value of a potential action.

We include the following quantities as features $\phi(s,a)$:
\begin{description}
\item[$P(C_a)$] The prior probability of the class that corresponds to the detector of action $a$
\item[$P(C_0|\mathbf{o}) \ldots P(C_K|\mathbf{o})$] The probabilities of all classes, conditioned on the current set of observations.
\item[$H(C_0|\mathbf{o}) \ldots H(C_K|\mathbf{o})$] The entropies of all classes, conditioned on the current set of observations.
\end{description}

Additionally, we include the mean and maximum entropies of all classes, the expected time of the action $c(a)$, and time features that represent the times until start and deadline, for a total of $F$ features.

We note that this setup is commonly used to solve Markov Decision Processes \cite{Sutton1998}.
There are two related limitations of MDPs when it comes to most systems of interesting complexity, however: the state has to be functionally approximated instead of exhaustively enumerated; and some aspects of the state are not observed, making the problem a Partially Observed MDP (POMDP), for which exact solution methods are intractable for all but rather small problems \cite{Roy2002}.

There isn't much to do about the necessity of approximating the state but design good features.
Our initial solution to the partial observability problem is to include uncertainty features into the feature representation to \emph{augment} the MDP \cite{Kwok2004}.

To formulate learning the policy as a single regression problem, we represent the features in block form, where $\phi(s,a_i)$ is a vector of size $F|\mathcal{A}|$, with all values set to $0$ except for the block corresponding to $a_i$.

As an illustration, we visualize the learned weights on these features in \autoref{fig:weights}, reshaped such that each row shows the weights learned for an action, in order.
The featurization for the $a_{gist}$ scene-context action, which concerns all classes, omits the first $P(C_a)$ feature.

@@

However, higher-order co-occurrences are not well represented in this form.
Additionally, updating $P(C_i|\mathbf{o})$ presents choices regarding independence assumptions between the classes.

We evaluate two approaches for updating probabilities: \emph{direct} and \emph{MRF}.
In the \emph{direct} method, $P(C_i|\mathbf{o}) = score(C_i)$ if $\mathbf{o}$ includes the observations for class $C_i$ and $P(C_i|\mathbf{o}) = P(C_i)$ otherwise.
This means that an observation of class $i$ does not influence the estimated probability of any class but $C_i$.
$score(C_i)$ for $a_{{det}_i}$ is obtained by training a probabilistic classifier on the detections output.
$score(C_i)$ for $a_{gist}$ is obtained by training probabilistic classifiers on the GIST feature, for all classes.

@@

The graphical model structure is set as fully-connected, but some classes are overwhelmingly unlikely to co-occurr in our dataset.
Accordingly, the graph edge weights are learned with $L_1$ regularization, which obtains the desired sparse structure \cite{Lee2006}.
All parameters of the model are trained on fully-observed data, and Loopy Belief Propagation inference is implemented with an open-source graphical model package \cite{Jaimovich2010}.
% As exact inference is generally intractable in this model, we use Loopy Belief Propagation; although it does not provide general convergence guarantees, it has been shown to work well empirically on similar tasks \cite{Desai2009}.

>>>>

As shown in Figure~\ref{fig:figure1}, the system is given two times: the setup time $T_s$ and deadline $T_d$.
We want to obtain the best possible answer if stopped at any given time between the setup time and the deadline.
A single-number metric that corresponds to this objective is the area captured under the curve between the start and deadline bounds, normalized by the total area.
We evaluate policies by this more robust metric and not simply by the final performance at deadline time for the same reason that Average Precision is used instead of a fixed Precision vs. Recall point in the conventional evaluations.

\subsection{Sequential Execution}
An \emph{open-loop} policy, such as the common classifier cascade \cite{Viola2001}, takes actions in a sequence that does not depend on observations received from previous actions.
In contrast, our goal is to learn a dynamic, or \emph{closed-loop}, policy, which would exploit the signal in scene and inter-object context for a maximally efficient path through the actions.

We refer to the information available to the decision process as the \emph{state} $s$.
The state includes the current estimate of the distribution over class presence variables $P(\mathbf{C}) = \{P(C_0), \ldots, P(C_K)\}$, where we write $P(C_k)$ to mean $P(C_k=1)$ (class $k$ is present in the image).

@@

The specific actions we consider in the following exposition are detector actions $a_{{det}_i}$, where ${det}_i$ is a detector class $C_i$, and a scene-level context action $a_{gist}$, which updates the probabilities of all classes.
Although we avoid this in the exposition, note that our system easily handles multiple detector actions per class.

\subsection{Selecting actions} \label{sec:value}
As our goal is to pick actions dynamically, we want a function $Q(s,a): S \times \mathcal{A} \mapsto \mathbb{R}$, where $S$ is the space of all possible states, to assign a value to a potential action $a \in \mathcal{A}$ given the current state $s$ of the decision process.
We can then define the policy $\pi$ as simply taking the action with the maximum value:
\begin{align}
\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} Q(s,a_i)
\end{align}

Although the action space $\mathcal{A}$ is manageable, the space of possible states $S$ is intractable, and we must use function approximation to represent $Q(s,a)$: a common technique in reinforcement learning \cite{Sutton1998}.
We featurize the state-action pair and assume linear structure:
\begin{align}
Q^\pi(s,a) = \theta_\pi^\top \phi(s,a)
\end{align}

The policy's performance at time $t$ is determined by all detections that are part of the set of observations $\mathbf{o}^j$ at the last state $s^j$ before $t$.
Recall that detector actions returns lists of detection hypotheses.
Therefore, the final AP vs. Time evaluation of an episode is a function $eval(h,T_s,T_d)$ of the history of execution $h=s^0,s^1,\dots,s^J$.
It is precisely the normalized area under the AP vs. Time curve between $T_s$ and $T_d$, as determined by the detections in $\mathbf{o}^j$ for all steps $j$ in the episode.

Note from~\autoref{fig:rewards} that this evaluation function is additive per action, as each action $a$ generates observations that may raise or lower the mean AP of the results so far ($\Delta ap$) and takes a certain time ($\Delta t$).
We can accordingly represent the final evaluation $eval(h,T_s,T_d)$ in terms of individual action rewards: $\sum_{j=0}^J R(s^j,a^j)$.

Specifically, as shown in Figure~\ref{fig:rewards}, we define the \emph{reward} of an action $a$ as
\begin{align}\label{eq:advanced}
R(s^j,a) = \Delta \text{ap} (t_T^j-\frac{1}{2}\Delta t)
\end{align}
where $t_T^j$ is the time left until $T_d$ at state $s^j$, and $\Delta t$ and $\Delta \text{ap}$ are the time taken and AP change produced by the action $a$.
(We do not account for $T_s$ here for clarity of exposition.)

\subsection{Learning the policy}
The expected value of the final evaluation can be written recursively in terms of the value function:
\begin{align} \label{eq:recursive_value}
Q^\pi(s^j,a) = \mathbb{E}_{s^{j+1}} [R(s^j,a) + \gamma Q^\pi(s^{j+1},\pi(s^{j+1}))]
\end{align}
where $\gamma \in [0,1]$ is the \emph{discount} value.

With $\gamma=0$, the value function is determined entirely by the immediate reward, and so only completely greedy policies can be learned.
With $\gamma=1$, the value function is determined by the correct expected rewards to the end of the episode.
However, a lower value of $\gamma$ mitigates the effects of increasing uncertainty regarding the state transitions over long episodes. %, and so the highest value is not necessarily best.
We set this meta-parameter of our approach through cross-validation, and find that a mid-level value ($0.4$) works best.

@@

We then learn the optimal policy by repeatedly gathering samples with the current policy, minimizing the error between the discounted reward to the end of the episode as predicted by our current $Q(s^j,a)$ and the actual values gathered, and updating the policy with the resulting weights.
%This is fitted Q-iteration, a variant of generalized policy iteration \cite{Ernst2005,Sutton1998}.

@@

To prevent overfitting to the training data, we use $L_2$-regularized regression.
We run $15$ iterations of accumulating samples by running $350$ episodes, starting with a baseline policy which will be described in \autoref{sec:evaluation}, and cross-validating the regularization parameter at each iteration.
Samples are not thrown away between iterations.

With pre-computed detections on the PASCAL VOC 2007 dataset, the training procedure takes about $4$ hours on an $8$-core \emph{Xeon E5620} machine.

\subsection{Feature representation}
Our policy is at its base determined by a linear function of the features of the state:
\begin{align}
\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \theta_\pi^\top \phi(s,a_i).
\end{align}

%Since we want to be able to learn a dynamic policy, the observations $\mathbf{o}$ that are part of the state $s$ should play a role in determining the value of a potential action.

We include the following quantities as features $\phi(s,a)$:

\begin{tabularx}{0.8\linewidth}{p{0.23\linewidth}p{0.69\linewidth}}
$P(C_a)$ & The prior probability of the class that corresponds to the detector of action $a$ (omitted for the scene-context action).\\
$P(C_0|\mathbf{o}) \ldots P(C_K|\mathbf{o})$ & The probabilities for all classes, conditioned on the current set of observations.\\
$H(C_0|\mathbf{o}) \ldots H(C_K|\mathbf{o})$ & The entropies for all classes, conditioned on the current set of observations. \\
\end{tabularx}

Additionally, we include the mean and maximum entropy of class presence and time features that represent the times until start and deadline, for a total of $F = 1+|K|*2+6$ features.

We note that this setup is commonly used to solve Markov Decision Processes \cite{Sutton1998}.
There are two related limitations of MDPs when it comes to most systems of interesting complexity, however: the state has to be functionally approximated instead of exhaustively enumerated; and some aspects of the state are not observed, making the problem a Partially Observed MDP (POMDP), for which exact solution methods are intractable for all but rather small problems \cite{Roy2002}.
Our initial solution to the problem of partial observability is to include features corresponding to our level of uncertainty into the feature representation, as in the technique of \emph{augmented} MDPs \cite{Kwok2004}.

To formulate learning the policy as a single regression problem, we represent the features in block form, where $\phi(s,a)$ is a vector of size $F|\mathcal{A}|$, with all values set to $0$ except for the $F$-sized block corresponding to $a$.

As an illustration, we visualize the learned weights on these features in \autoref{fig:weights}, reshaped such that each row shows the weights learned for an action, with the top row representing the scene context action and then next $20$ rows corresponding to the PASCAL VOC class detector actions.

@@

However, higher-order co-occurrences are not well represented in this form.
Additionally, updating $P(C_i|\mathbf{o})$ presents choices regarding independence assumptions between the classes.
We evaluate two approaches for updating probabilities: \emph{direct} and \emph{MRF}.

In the \emph{direct} method, $P(C_i|\mathbf{o}) = score(C_i)$ if $\mathbf{o}$ includes the observations for class $C_i$ and $P(C_i|\mathbf{o}) = P(C_i)$ otherwise.
This means that an observation of class $i$ does not influence the estimated probability of any class but $C_i$.

@@

The graphical model structure is set as fully-connected, but some classes almost never co-occurr in our dataset.
Accordingly, the edge weights are learned with $L_1$ regularization, which obtains a sparse structure \cite{Lee2006}.
All parameters of the model are trained on fully-observed data, and Loopy Belief Propagation inference is implemented with an open-source graphical model package \cite{Jaimovich2010}.

An implementation detail: $score(C_i)$ for $a_{{det}_i}$ is obtained by training a probabilistic classifier on the list of detections, featurized by the top few confidence scores and the total number of detections.
Similarly, $score(C_i)$ for $a_{gist}$ is obtained by training probabilistic classifiers on the GIST feature, for all classes.
